Amphistomy increases leaf photosynthesis more in coastal than montane plants of Hawaiian ʻilima (\emph{Sida fallax})

### 2023-12-19 ms/ms.qmd
added Zenodo link to ms
fixed a couple typos

where $g_\mathrm{smax,lower}$ and $g_\mathrm{smax,upper}$ are maximum stomatal conductance to water vapor at $T_\mathrm{leaf} = 25 ^ \circ$ C on the lower and upper surface, respectively. The maximum stomatal conductance was calculated from stomatal density and length, assuming that stomata are fully open, following @sack_developmental_2016:

@@

Three limitations of this study are the small sample size, experimental design that precludes distinguishing genetic from environmental differences in leaf traits, and potentially confounding effects of other environmental differences besides light environment. Understanding the mechanistic basis of higher $\mathrm{AA}$ in sun leaves would require much larger sample sizes. Sun leaves tend to be thicker, more densely packed with mesophyll cells, and have greater photosynthetic capacity and higher stomatal conductance, among other traits [@lambers_plant_2008]. Each of these factors and others potentially modulate $\mathrm{AA}$. Quantifying the contribution of all these factors requires larger samples and additional measurements that are beyond the scope of this study, but exciting avenues for future research on leaf structure-function relations. Although many morphological traits that distinguish coastal and montane ʻilima populations persist in a common environment [@yorkston_interfertility_2006], we cannot distinguish between genetic effects and plastic responses to habitat as causes of difference in $\mathrm{AA}$ because we measured naturally occurring plants *in situ*. While disentangling genetic and plastic contributions is not necessarily important for understanding the distribution of amphistomatous leaves, it would be insightful to know about genetic and environmental contributions to trait variation. A reciprocal transplant would be able to determine the genetic and environmental contributions, as well their interaction, to trait variance in nature. However, reciprocal transplants cannot control for other differences between coastal and montane habitats besides vegetation height, such as temperature and precipitaiton. Experimental studies in controlled environments will be necessary to isolate the effects of light quantity and quality on $\mathrm{AA}$.

>>>>

where $g_\mathrm{smax,lower}$ and $g_\mathrm{smax,upper}$ are the maximum stomatal conductance to water vapor at $T_\mathrm{leaf} = 25 ^ \circ$ C on the lower and upper surface, respectively. The maximum stomatal conductance was calculated from stomatal density and length, assuming that stomata are fully open, following @sack_developmental_2016:

@@

Three limitations of this study are the small sample size, experimental design that precludes distinguishing genetic from environmental differences in leaf traits, and potentially confounding effects of other environmental differences besides light environment. Understanding the mechanistic basis of higher $\mathrm{AA}$ in sun leaves would require much larger sample sizes. Sun leaves tend to be thicker, more densely packed with mesophyll cells, and have greater photosynthetic capacity and higher stomatal conductance, among other traits [@lambers_plant_2008]. Each of these factors and others potentially modulate $\mathrm{AA}$. Quantifying the contribution of all these factors requires larger samples and additional measurements that are beyond the scope of this study, but exciting avenues for future research on leaf structure-function relations. Although many morphological traits that distinguish coastal and montane ʻilima populations persist in a common environment [@yorkston_interfertility_2006], we cannot distinguish between genetic effects and plastic responses to habitat as causes of difference in $\mathrm{AA}$ because we measured naturally occurring plants *in situ*. While disentangling genetic and plastic contributions is not necessarily important for understanding the distribution of amphistomatous leaves, it would be insightful to know about genetic and environmental contributions to trait variation. A reciprocal transplant would be able to determine the genetic and environmental contributions, as well their interaction, to trait variance in nature. However, reciprocal transplants cannot control for other differences between coastal and montane habitats besides vegetation height, such as temperature and precipitation. Experimental studies in controlled environments will be necessary to isolate the effects of light quantity and quality on $\mathrm{AA}$.

### 2023-12-19 ms/ms.tex
fixed a couple typos

% toggle line numbers
\usepackage{lineno}
\linenumbers

@@

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, sharp corners, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

To estimate \(\hat{\beta}\)'s from the \(A \textendash g_\text{sw}\)
curve for each leaf, we fit Bayesian regressions using the \emph{R}
package \textbf{brms} version 2.20.4 (Bürkner, 2017) with MCMC sampling
in \emph{Stan} (Stan Development Team, 2023). We used CmdStan version
2.33.1 and \textbf{cmdstanr} version 0.6.1 (Gabry et al., 2023) to
interface with \emph{R} version 4.3.1 (R Core Team, 2023). We sampled
the posterior distribution from 4 chains with 1000 iterations each after
1000 warmup iterations per chain. We estimated parameters and confidence
intervals as the median and 95\% quantile intervals of the posterior,
respectively. The key prediction is that
\(\mathrm{AA}_\text{coastal} > \mathrm{AA}_\text{montane}\), meaning the
95\% confidence intervals of
\(\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane}\) should be
positive and not encompass 0.

@@

where \(g_\mathrm{smax,lower}\) and \(g_\mathrm{smax,upper}\) are
maximum stomatal conductance to water vapor at
\(T_\mathrm{leaf} = 25 ^ \circ\) C on the lower and upper surface,
respectively. The maximum stomatal conductance was calculated from
stomatal density and length, assuming that stomata are fully open,
following Sack and Buckley (2016):

@@

Three limitations of this study are the small sample size, experimental
design that precludes distinguishing genetic from environmental
differences in leaf traits, and potentially confounding effects of other
environmental differences besides light environment. Understanding the
mechanistic basis of higher \(\mathrm{AA}\) in sun leaves would require
much larger sample sizes. Sun leaves tend to be thicker, more densely
packed with mesophyll cells, and have greater photosynthetic capacity
and higher stomatal conductance, among other traits (Lambers et al.,
2008). Each of these factors and others potentially modulate
\(\mathrm{AA}\). Quantifying the contribution of all these factors
requires larger samples and additional measurements that are beyond the
scope of this study, but exciting avenues for future research on leaf
structure-function relations. Although many morphological traits that
distinguish coastal and montane ʻilima populations persist in a common
environment (Yorkston and Daehler, 2006), we cannot distinguish between
genetic effects and plastic responses to habitat as causes of difference
in \(\mathrm{AA}\) because we measured naturally occurring plants
\emph{in situ}. While disentangling genetic and plastic contributions is
not necessarily important for understanding the distribution of
amphistomatous leaves, it would be insightful to know about genetic and
environmental contributions to trait variation. A reciprocal transplant
would be able to determine the genetic and environmental contributions,
as well their interaction, to trait variance in nature. However,
reciprocal transplants cannot control for other differences between
coastal and montane habitats besides vegetation height, such as
temperature and precipitaiton. Experimental studies in controlled
environments will be necessary to isolate the effects of light quantity
and quality on \(\mathrm{AA}\).

>>>>

% toggle line numbers
\usepackage{lineno}
% \linenumbers

@@

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, sharp corners, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

@@

To estimate \(\hat{\beta}\)'s from the \(A \textendash g_\text{sw}\)
curve for each leaf, we fit Bayesian regressions using the \emph{R}
package \textbf{brms} version 2.20.4 (Bürkner, 2017) with MCMC sampling
in \emph{Stan} (Stan Development Team, 2023). We used CmdStan version
2.33.1 and \textbf{cmdstanr} version 0.6.1 (Gabry et al., 2023) to
interface with \emph{R} version 4.3.2 (R Core Team, 2023). We sampled
the posterior distribution from 4 chains with 1000 iterations each after
1000 warmup iterations per chain. We estimated parameters and confidence
intervals as the median and 95\% quantile intervals of the posterior,
respectively. The key prediction is that
\(\mathrm{AA}_\text{coastal} > \mathrm{AA}_\text{montane}\), meaning the
95\% confidence intervals of
\(\mathrm{AA}_\text{coastal} - \mathrm{AA}_\text{montane}\) should be
positive and not encompass 0.

@@

where \(g_\mathrm{smax,lower}\) and \(g_\mathrm{smax,upper}\) are the
maximum stomatal conductance to water vapor at
\(T_\mathrm{leaf} = 25 ^ \circ\) C on the lower and upper surface,
respectively. The maximum stomatal conductance was calculated from
stomatal density and length, assuming that stomata are fully open,
following Sack and Buckley (2016):

@@

Three limitations of this study are the small sample size, experimental
design that precludes distinguishing genetic from environmental
differences in leaf traits, and potentially confounding effects of other
environmental differences besides light environment. Understanding the
mechanistic basis of higher \(\mathrm{AA}\) in sun leaves would require
much larger sample sizes. Sun leaves tend to be thicker, more densely
packed with mesophyll cells, and have greater photosynthetic capacity
and higher stomatal conductance, among other traits (Lambers et al.,
2008). Each of these factors and others potentially modulate
\(\mathrm{AA}\). Quantifying the contribution of all these factors
requires larger samples and additional measurements that are beyond the
scope of this study, but exciting avenues for future research on leaf
structure-function relations. Although many morphological traits that
distinguish coastal and montane ʻilima populations persist in a common
environment (Yorkston and Daehler, 2006), we cannot distinguish between
genetic effects and plastic responses to habitat as causes of difference
in \(\mathrm{AA}\) because we measured naturally occurring plants
\emph{in situ}. While disentangling genetic and plastic contributions is
not necessarily important for understanding the distribution of
amphistomatous leaves, it would be insightful to know about genetic and
environmental contributions to trait variation. A reciprocal transplant
would be able to determine the genetic and environmental contributions,
as well their interaction, to trait variance in nature. However,
reciprocal transplants cannot control for other differences between
coastal and montane habitats besides vegetation height, such as
temperature and precipitation. Experimental studies in controlled
environments will be necessary to isolate the effects of light quantity
and quality on \(\mathrm{AA}\).
